Wireless service accounting based on Ethereum state channels - Master Thesis

### 2019-10-06 Paper.tex

\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
\usepackage{lastpage}
\usepackage{footnote}
\usepackage{caption}
\definecolor{Orange}{rgb}{0.241,0.349,0.090}
\usepackage{bbding}
\usepackage{shortvrb}
%% \usepackage{draftwatermark}
%% \SetWatermarkText{Esborrany}
%% \SetWatermarkScale{0.5}
%% \usepackage{tocloft}
% \usepackage{hyperref}
% \usepackage{bookmark}
\usepackage[titletoc]{appendix}

@@

\begin{abstract}
  \begin{center}
    This document contains authors' ``\textit{Blockchain
      technologies master}''\footnote{UPC School
      of Professional \& Executive Development}
    final thesis. The project described in the document
    consists on a technological proof
    of concept of an internet access service consumption
    using a state channel smart contract as
    mean of payment. After a brief
    introduction, there's a first motivation of
    the use case included;
    in the following sections
    there's a detailed explanation of
    the architecture, implementation and
    security considerations;
    after that it is described a more
    general approach about using blockchain technologies
    and state channels in particular to provide
    mechanisms to reduce advanced wireless
    (i. e. 5G) services market friction between neutral host
    providers, virtual operators and operators
    licensing spectrum bands to nations' governments;
    guarantee transparency to consumers,
    and provide a fully GDPR compliant
    schema using a self-sovereign identity system.
  \end{center}
\end{abstract}

@@

@@

\section{Technological Proof of Concept description}

@@

@@

\begin{itemize}
\item[] 1) The user associates the laptop with the WiFi hotspot.
\item[] 2) Once connected is redirected to the captive portal front-end.
\item[] 3) Front provisions MetaMask as Web3 provider.
\item[] 4) Front-end uses MetaMaskto allow the user
  to send value to the contract
\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
  on port 8545.
\item[] 7) Once transaction is confirmed, the back-end iptables controller
  start forwarding user's traffic.
\end{itemize}

@@

@@

\subsection{Implementation} \label{ch:implementation}

>>>>

\usepackage[final]{microtype} % https://tex.stackexchange.com/questions/75140/is-it-possible-to-make-latex-mark-overfull-boxes-in-the-output#comment382776_75142
\usepackage{lastpage}
\usepackage{footnote}
\usepackage{caption}
\definecolor{Orange}{rgb}{0.241,0.349,0.090}
% \usepackage{bbding}
\usepackage{shortvrb}
%% \usepackage{draftwatermark}
%% \SetWatermarkText{Esborrany}
%% \SetWatermarkScale{0.5}
%% \usepackage{tocloft}
% \usepackage{hyperref}
% \usepackage{bookmark}
\usepackage[titletoc]{appendix}

@@

\begin{abstract}
  \begin{center}
    This document contains authors' ``\textit{Blockchain
      technologies master}''\footnote{UPC School
      of Professional \& Executive Development}
    final thesis. The project described in the document
    consists on a technological proof
    of concept of an internet access service consumption
    using a state channel smart contract as
    mean of payment. After a brief
    introduction, there's a first motivation of
    the use case included;
    in the following sections
    there's a detailed explanation of
    the architecture, implementation and
    security considerations;
    after that it is described a more
    general approach about using blockchain technologies
    and state channels in particular to provide
    mechanisms to reduce advanced wireless
    (i. e. 5G) services market friction between neutral host
    providers, virtual operators and operators
    purchasing licenses of spectrum bands to nations' governments;
    guarantee transparency to consumers,
    and provide a fully GDPR compliant
    schema relying on a self-sovereign identity system.
  \end{center}
\end{abstract}

@@

Ethereum
blockchain is proposed as a first optimal candidate platform
to support this project PoC attending to
following reasons:

\begin{itemize}
\item Is the most used and mature 
turing-complete-programmable blockchain technology.
\item Support for high quality and audited 
libraries in Solidity (most used smart contract language in Ethereum).
\item Developments on Ethereum can be deployed 
over public (main or testnets) or private/permissioned networks.
\item Unlike Bitcoin, Ethereum supports miner fees using gas
instead of directly ETH; this can mitigate the effects of
volatile crypto-currency price by adjusting gas price depending
on ETH value.
\end{itemize}

@@

\section{Proof of Concept description}

@@

\subsection{General work-flow}

@@

\begin{itemize}
\item[] 1) The user associates the laptop with the WiFi hotspot.
\item[] 2) Once connected is redirected to the captive portal front-end.
\item[] 3) Front provisions MetaMask as Web3 provider.
\item[] 4) Front-end uses MetaMaskto allow the user
  to send value to the contract opening a \textit{state channel} 
(\textit{see section below}).
\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
  on port 8545.
\item[] 7) Once transaction is confirmed, the back-end iptables controller
  start forwarding user's traffic.
\end{itemize}

@@

In the recent years blockchain technologies
and in particular smart contracts have made possible
an innovative way to rule business and
workflows between organizations.

One of the main application of smart contracts
consist on modeling a scarce resource representing
it using fungible (or not) tokens; in addition
to this, smart contracts can be programmed
to suit any possible business intelligence
according to rules well known and accepted
in advance by the parties transacting.

In the case of 5G technologies and due to the
requirement that different operators shall be able
to dynamically share the same resources (i.e. neutral host
infrastructure). At the same time in order to manage
service operation efficiently through its whole
life-cycle there's the need to automate as much as
possible all the operations; this requirements
perfectly fit smart contract capabilities, in order
to model and operate over a defined set of rules
supporting payments, escrows, rewards or penalties
depending on different conditions.

@@

% \subsection{Implementation} \label{ch:implementation}
